---
title: "R Chart Example with Multiple Libraries"
author: "Jane Adams"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    fig_width: 12
    fig_height: 4
---

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)
library(lattice)
library(gridExtra)
library(ggplot2)
library(Cairo)
```

# Create Random Data to Visualize

```{r data-generation}
set.seed(0)
N <- 100
D <- c('A', 'B', 'C')
df <- tibble(
  c1 = runif(N),
  c2 = runif(N),
  d1 = sample(D, N, replace = TRUE),
  d2 = sample(D, N, replace = TRUE)
)
agg_df <- df %>% group_by(d1) %>% summarise(c1 = mean(c1))
pivot <- table(df$d1, df$d2)
```

# ggplot2 Visualizations

```{r ggplot2-plots, fig.width=12, fig.height=4}
if (!dir.exists("plots")) dir.create("plots")
p1 <- ggplot(df, aes(x = c1, y = c2, color = d1)) +
  geom_point(size = 3, alpha = 0.8) +
  ggtitle("Continuous-Continuous") +
  theme_minimal()

p2 <- ggplot(agg_df, aes(x = d1, y = c1, fill = d1)) +
  geom_col() +
  ggtitle("Continuous-Discrete") +
  theme_minimal()

p3 <- ggplot(as.data.frame(pivot), aes(Var1, Var2, fill = Freq)) +
  geom_tile() +
  scale_fill_viridis_c() +
  ggtitle("Discrete-Discrete") +
  labs(x = "d1", y = "d2") +
  theme_minimal()

gridExtra::grid.arrange(p1, p2, p3, ncol = 3)
```

# Save ggplot2 plots to SVG using Cairo
```{r save-ggplot2-svg, echo=FALSE, message=FALSE, warning=FALSE}
Cairo::CairoSVG(file = "plots/ggplot.svg", width = 12, height = 4)
print(gridExtra::grid.arrange(p1, p2, p3, ncol = 3))
dev.off()
message("SVG file saved to plots/ggplot.svg")
```

# Save ggplot2 plots to SVG using svglite
```{r save-ggplot2-svg-svglite, echo=FALSE, message=FALSE, warning=FALSE}
svglite::svglite("plots/ggplot.svg", width = 12, height = 4)
gridExtra::grid.arrange(p1, p2, p3, ncol = 3)
dev.off()
```

# Plotly Visualizations

```{r plotly-plots, eval=interactive()}
p1 <- plot_ly(df, x = ~c1, y = ~c2, color = ~d1, type = 'scatter', mode = 'markers') %>%
  layout(title = "Continuous-Continuous")
p2 <- plot_ly(agg_df, x = ~d1, y = ~c1, color = ~d1, type = 'bar') %>%
  layout(title = "Continuous-Discrete")
p3 <- plot_ly(z = ~pivot, x = colnames(pivot), y = rownames(pivot), type = "heatmap", colors = "viridis") %>%
  layout(title = "Discrete-Discrete")

subplot(p1, p2, p3, nrows = 1, margin = 0.05, titleX = TRUE, titleY = TRUE) %>%
  layout(width = 1400, height = 400)
```

# Save plotly plots to SVG using Kaleido
```{r save-plotly-svg, echo=FALSE, message=FALSE, warning=FALSE}
subplot_obj <- subplot(p1, p2, p3, nrows = 1, margin = 0.05, titleX = TRUE, titleY = TRUE)
plotly::save_image(subplot_obj, file = "plots/R-plotly.svg", engine = "kaleido", width = 1200, height = 400)
message("SVG file saved to plots/R-plotly.svg")
```

# Lattice Visualizations

```{r lattice-plots, fig.width=12, fig.height=4}
p1 <- xyplot(c2 ~ c1, data = df, groups = d1, auto.key = TRUE, main = "Continuous-Continuous",
             par.settings = list(background = list(col = "white")))
p2 <- barchart(c1 ~ d1, data = agg_df, groups = d1, main = "Continuous-Discrete",
               par.settings = list(background = list(col = "white")))
p3 <- levelplot(pivot, main = "Discrete-Discrete", xlab = "d2", ylab = "d1",
                col.regions = viridisLite::viridis(100),
                par.settings = list(background = list(col = "white")))
gridExtra::grid.arrange(p1, p2, p3, ncol = 3)
```

# Save lattice plots to SVG using Cairo
```{r save-lattice-svg, echo=FALSE, message=FALSE, warning=FALSE}
Cairo::CairoSVG(file = "plots/lattice.svg", width = 12, height = 4)
p1 <- xyplot(c2 ~ c1, data = df, groups = d1, auto.key = TRUE, main = "Continuous-Continuous",
             par.settings = list(background = list(col = "white")))
p2 <- barchart(c1 ~ d1, data = agg_df, groups = d1, main = "Continuous-Discrete",
               par.settings = list(background = list(col = "white")))
p3 <- levelplot(pivot, main = "Discrete-Discrete", xlab = "d2", ylab = "d1",
                col.regions = viridisLite::viridis(100),
                par.settings = list(background = list(col = "white")))
gridExtra::grid.arrange(p1, p2, p3, ncol = 3)
dev.off()
message("SVG file saved to plots/lattice.svg")
```

# Base R Visualizations

```{r base-plots, fig.width=12, fig.height=4}
par(bg = "white")
par(mfrow = c(1, 3))
# Continuous-Continuous
plot(df$c1, df$c2, col = as.numeric(factor(df$d1)), pch = 19, main = "Continuous-Continuous", xlab = "c1", ylab = "c2")
legend("topright", legend = levels(factor(df$d1)), col = 1:3, pch = 19)
# Continuous-Discrete
barplot(agg_df$c1, names.arg = agg_df$d1, col = 1:3, main = "Continuous-Discrete", ylab = "mean(c1)")
# Discrete-Discrete
image(1:3, 1:3, t(pivot)[,3:1], col = viridisLite::viridis(100), axes = FALSE, main = "Discrete-Discrete", xlab = "d2", ylab = "d1")
axis(1, at = 1:3, labels = colnames(pivot))
axis(2, at = 1:3, labels = rev(rownames(pivot)))
par(mfrow = c(1, 1))
```

# Save base R plots to SVG using Cairo
```{r save-base-svg, echo=FALSE, message=FALSE, warning=FALSE}
Cairo::CairoSVG(file = "plots/base.svg", width = 12, height = 4, bg = "white")
par(bg = "white")
par(mfrow = c(1, 3))
# Continuous-Continuous
plot(df$c1, df$c2, col = as.numeric(factor(df$d1)), pch = 19, main = "Continuous-Continuous", xlab = "c1", ylab = "c2")
legend("topright", legend = levels(factor(df$d1)), col = 1:3, pch = 19)
# Continuous-Discrete
barplot(agg_df$c1, names.arg = agg_df$d1, col = 1:3, main = "Continuous-Discrete", ylab = "mean(c1)")
# Discrete-Discrete
image(1:3, 1:3, t(pivot)[,3:1], col = viridisLite::viridis(100), axes = FALSE, main = "Discrete-Discrete", xlab = "d2", ylab = "d1")
axis(1, at = 1:3, labels = colnames(pivot))
axis(2, at = 1:3, labels = rev(rownames(pivot)))
par(mfrow = c(1, 1))
dev.off()
message("SVG file saved to plots/base.svg")
```

---

This document demonstrates how to create and compare charts using multiple R libraries: ggplot2, plotly, lattice, and base R. Each section generates three types of plots: continuous-continuous, continuous-discrete, and discrete-discrete.
